---
title: "Rationale for MASSExtra"
author: "Bill Venables"
date: "`r Sys.Date()`"
output: 
  html_vignette: NULL
    pdf_document:
    includes:
      in_header: header.tex
vignette: >
  %\VignetteIndexEntry{Rationale for MASSExtra}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = ""
)
```

# Preamble

This extension package to the classical `MASS` package, whose origins go back to nearly 30
years, comes about for a number of reasons.  

Firstly, in my teaching I found I was using some of the old functions in the package with
consistently different argument settings to the defaults.  I was also interested in supplying
various convenience extensions that simplified teaching and including various tweaks to
improve the interface.  Examples follow below.

Secondly, I wanted to provide a few functions that were mainly useful as programming examples.
For example, the function `zs` and its allies `zu`, `zq` and `zr` are mainly alternatives to
`base::scale`, but they can be used to show how to write functions that can be used in fitting
models in such a way that they work as they should when the fitted model object is used for
prediction with new data.

Finally, there is the perennial `select` problem.  When `MASS` is used with other packages,
such as `dplyr` the `select` function can easly be masked, causing confusion with users.
`MASS::select` is rarely used, but `dplyr::select` is fundamental.  There are standard ways
of managing this kind of masking, but what we have done in `MASSExtra` is to export the
more common functions used from `MASS` along with the extensions, in such a way that users
will not need to have `MASS` attached to the search path at all, and hence masking is unlikely.



```{r setup, fig.height = 5, fig.width = 10, out.width="100%"}
library(MASSExtra)
par(mfrow = c(1,2))
mod0 <- lm(MPG.city ~ Weight, Cars93)
boxcox(mod0)  ## MASS
box_cox(mod0) ## MASSExtra tweak
```
